Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-08 paper/paper.tex
fixed macro, moved -ve sign in (21)
finished empirical bayes

%% drafting macro
\newcommand{\important}[1]{\textbf{\colorbox{red}{#1}}}
\newcommand{\attn}[2]{\textbf{\color{red} #2~\textsuperscript{\textit{[#1]}}}}
\newcommand{\verify}[1]{\attn{verify}{#1}}
\newcommand{\rewrite}[1]{\attn{rewrite}{#1}}
\newcommand{\todo}[1]{{\color{blue} [TODO]} \important{#1}}

@@

We consider two priors for Bayesian dropout: log-uniform and $\cplx$-Gaussian empirical Bayes,
\cite{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}. For a 
fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$, the
divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q(\omega) \| \pi(\omega))
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij} \mathbb{H}(q(\omega_{ij}))
      + \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We use approximation \eqref{eq:c-gauss-vi} and omit subscripts ${ij}$ below for brevity.

@@

Note that for $\mu \neq 0$ from \eqref{eq:cn-affine} we have $
  \mathcal{N}^{\cplx}(\mu, \alpha \lvert \mu \rvert^2, 0)

@@

@@

@@

$ (Euler's constant) as $x\to 0-$ and $

@@

$ on $x \in (-\infty, -1]$.

@@

For this model we use an empirical Bayes prior. We assume that our initial belief is that
parameters are independent and each one is proper $\cplx$-Gaussian with its own precision: $
  \pi(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element KL-divergence term in the ELBO is computed thus:
\begin{equation}  \label{eq:emp-bayes-kl-div}
KL(q \| \pi_\tau)
  = - 1 - \log{(\tau \sigma^2)}
    + \tau \bigl(
      \sigma^2 + \lvert \mu \rvert^2
    \bigr)
  \,,
\end{equation}
which follows from the KL-divergence expression for two multivariate Gaussians
$$
% \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
%   =
KL(q_1\| q_2)
  =
  % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
  - \tfrac12 \log \det{e I}
  + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
  + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
  + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
  \,. $$

Somewhat counterintuitively, in empirical Bayes we optimize the hyperparameters of the
prior on the observed dataset. Thus the lower bound \eqref{eq:elbo} is optimized w.r.t.
$\tau$ of each weight's prior. \eqref{eq:emp-bayes-kl-div} is optimized (minimized) at
$\tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}$, yielding the following penalty in
ELBO:
\begin{equation}  \label{eq:emp-bayes-opt-kl}
KL(q \| \pi_{\tau^\ast})
  % = - 1 - \log{({\tau^\ast} \sigma^2)}
  %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
  % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
  = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
  % = \log{\bigl(1 + \tfrac1\alpha \bigr)}
  \,,
\end{equation}
which exactly replicates the term, derived in \cite{kharitonov_variational_2018}.

>>>>

%% drafting macro
\newcommand{\important}[1]{\textbf{\!\colorbox{red}{#1}\!}}
\newcommand{\attn}[2]{\textbf{\color{red} #2~\textsuperscript{\textit{[#1]}}}}
\newcommand{\verify}[1]{\textit{\!\colorbox{red}{#1}\!}}
\newcommand{\rewrite}[1]{\attn{rewrite}{#1}}
\newcommand{\todo}[1]{{\color{blue} [TODO]} \important{#1}}

@@

We consider two priors for Bayesian dropout: log-uniform and $\cplx$-Gaussian empirical Bayes,
\cite{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}. For a 
fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$, the
divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q(\omega) \| \pi(\omega))
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = \sum_{ij} \bigl(
        - \mathbb{H}(q(\omega_{ij}))
        - \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})} \log{\pi(\omega_{ij})}
      \bigr)
    \,.
\end{equation}
We use approximation \eqref{eq:c-gauss-vi} and omit subscripts ${ij}$ below for brevity.

@@

\eqref{eq:cn-affine} implies that $
  \mathcal{N}^{\cplx}(\mu, \sigma^2, 0)

@@

$ for $\mu \neq 0$ and $
  \sigma^2 = \alpha \lvert \mu \rvert^2

@@

  % as x to 0 from below

@@

$ as $x\to 0$ ($\gamma$ is Euler's constant) and $

@@

$ for $x \leq -1$.

@@

Let's consider fully factorized proper $\cplx$-gaussian prior $
  \pi_\tau(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element divergence term in \eqref{eq:elbo-general-kl-div} is
\begin{equation}  \label{eq:emp-bayes-kl-div}
  KL(q \| \pi_\tau)
    = - 1 - \log{(\tau \sigma^2)}
      + \tau \bigl(
        \sigma^2 + \lvert \mu \rvert^2
      \bigr)
    \,.
\end{equation}
%
% This follows from the KL-divergence expression for two multivariate Gaussians
% $$
% % \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
% %   =
% KL(q_1\| q_2)
%   =
%   % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
%   - \tfrac12 \log \det{e I}
%   + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
%   + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
%   + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
%   \,. $$
%
In empirical Bayes the prior adapts to the observed data, i.e. \eqref{eq:elbo} is optimized
w.r.t. $\tau$ of each weight's prior. Since divergence sits in \eqref{eq:elbo} with negative
sign, optimal $
  \tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}
$ is found by minimizing \eqref{eq:emp-bayes-kl-div}, which gives
\begin{equation}  \label{eq:emp-bayes-opt-kl}
  KL(q \| \pi_{\tau^\ast})
    % = - 1 - \log{({\tau^\ast} \sigma^2)}
    %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
    % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
    = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
    % = \log{\bigl(1 + \tfrac1\alpha \bigr)}
    \,.
\end{equation}
